---
title: "exercise-10"
author: "Minsoo Joo"
date: "2025-04-04"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
```

```{r}
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/AVONETdataset1.csv"
d<-read_csv(f, col_names=TRUE)
d <- d %>% 
  select(Species1, Family1, Order1,
         Beak.Length_Culmen, Beak.Width, Beak.Depth,
         Tarsus.Length, Wing.Length, Tail.Length, Mass,
         Habitat, Migration, Trophic.Level, Trophic.Niche,
         Min.Latitude, Max.Latitude, Centroid.Latitude,
         Primary.Lifestyle, Range.Size)

skim(d)
```

Categorical variables: Species1, Family1, Order1, Habitat, Trophic.Level, Trophic.Niche, Primary.Lifestyle

Numerical variables: all other variables

# Challenge 1

## Step 1

```{r}
d <- d %>% 
  filter(!is.na(Trophic.Level), !is.na(Migration))
# Convert Migration to factor
d <- d %>%
  mutate(Migration = as.factor(Migration))

d %>% 
  ggplot(aes(x = Trophic.Level, y = log(Mass))) +
  geom_boxplot()

d %>% 
  ggplot(aes(x = Migration, y = log(Mass))) +
  geom_boxplot()


```

## Step 2

Both are significantly associated with log(Mass).
(p-value \< 0.05)

Both Migration2 and Migration3 differ significantly from Migration1.
(very small p-values)

Migration1 is the reference level.

```{r}
summary(lm(log(Mass) ~ Trophic.Level, data=d))
summary(lm(log(Mass) ~ Migration, data=d))
```

All three categories are significantly different from each other.

```{r}
summary(lm(log(Mass) ~ relevel(d$Migration, ref = "2"), data = d))
```

## Step 3

All migration groups differ significantly from one another.

```{r}
m <- aov(log(Mass) ~ Migration, data = d)
posthoc <- TukeyHSD(m, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc 
```

## Step 4

```{r}
library(infer)

d <- d %>%
  mutate(logMass = log(Mass))

obs_f <- d %>%
  specify(logMass ~ Trophic.Level) %>%
  calculate(stat = "F")

# Null distribution
null_dist <- d %>%
  specify(logMass ~ Trophic.Level) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "F")

# Empirical p-value
get_p_value(null_dist, obs_stat = obs_f, direction = "greater")
```

# Challenge 2

## Step 1

```{r}
relBeak <- lm(log(d$Beak.Length_Culmen)~log(d$Mass), data=d)
relTarsus <- lm(log(d$Tarsus.Length)~log(d$Mass), data=d)
d <- d %>% 
  mutate(relBeak = relBeak$residuals,
         relTarsus = relTarsus$residuals)
```

## Step 2

```{r}
ggplot(d, aes(x = Primary.Lifestyle, y = relTarsus)) +
  geom_boxplot()

ggplot(d, aes(x = Trophic.Niche, y = relBeak)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Step 3

Migration is significantly associated with log(Range.Size).

Migration behavior explains approximately 8.7% of the variance in log-transformed geographic range size.

```{r}
d <- d %>%
  filter(!is.na(Migration))

# Right-skewed
ggplot(d, aes(x = Range.Size)) +
  geom_histogram(bins = 30)

# It needs to be log-transformed
d <- d %>%
  mutate(logRS = log(Range.Size))

m <- lm(logRS ~ Migration, data = d)
summary(m)
```

Both Migration2 and Migration3 are significantly different from the reference (Migration1).

All three migration categories are significantly different from each other.
(p-value \< 0)

```{r}
m2 <- lm(logRS ~ relevel(d$Migration, ref = "2"), data = d)
summary(m2)

m3 <- aov(log(Mass) ~ Migration, data = d)
posthoc <- TukeyHSD(m3, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc 
```

## Step 4

```{r}
d_p <- d %>% 
  filter(Order1 == "Passeriformes")
```

```{r}
ggplot(d_p, aes(x = Primary.Lifestyle, y = relBeak)) +
  geom_boxplot()

ggplot(d_p, aes(x = Trophic.Level, y = relBeak)) +
  geom_boxplot()
```

Primary.Lifestyle is strongly associated with relative beak length.

And there is a significant difference between groups.

```{r}
m3 <- lm(relBeak ~ Primary.Lifestyle, data = d_p)
summary(m3)
anova(m3)
```

Trophic.Level is strongly associated with relative beak length.

And there is a significant difference between groups.

```{r}
m4 <- lm(relBeak ~ Trophic.Level, data = d_p)
summary(m4)
anova(m4)
```

```{r}
unique(d_p$Primary.Lifestyle)
unique(d_p$Trophic.Level)
```

## Step 5

Two-way ANOVA showed that both Primary.Lifestyle and Trophic.Level are significantly associated with relative beak length.

Primary.Lifestyle: All three show significantly longer relative beaks than Aerial species.

Trophic.Level: Herbivore clearly differ from carnivore, while omnivore is similar.

```{r}
m5 <- lm(relBeak ~ Primary.Lifestyle + Trophic.Level, data = d_p)
summary(m5)
anova(m5)
```

## Step 6

Relative beak length is significantly influenced by both primary lifestyle and trophic level, and their interaction indicates that the effect of one depends on the other.

```{r}
m6 <- lm(relBeak ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level, data = d_p)
summary(m6)
anova(m6)
```

## Step 7

```{r}
interaction.plot(x.factor = d_p$Primary.Lifestyle, xlab = "Primary.Lifestyle",
                 trace.factor = d_p$Trophic.Level, trace.label = "Trophic.Level",
                 response = d_p$relBeak, fun = base::mean, ylab = "Mean Relative Beak Length")
```

## Step 8

Less than 2.
(the assumption may not be violated)

```{r}
# Trophic.Level
d_p %>%
  group_by(Trophic.Level) %>%
  summarise(n = n(), sd_relBeak = sd(relBeak)) %>%
  summarise(ratio = max(sd_relBeak) / min(sd_relBeak))

# Primary.Lifestyle
d_p %>%
  group_by(Primary.Lifestyle) %>%
  summarise(n = n(), sd_relBeak = sd(relBeak)) %>%
  summarise(ratio = max(sd_relBeak) / min(sd_relBeak))
```

Residuals were approximately normally distributed across most groups.

Minor deviations were observed, but overall the normality assumption appears reasonably met.

```{r}
mod <- lm(relBeak ~ Trophic.Level + Primary.Lifestyle, data = d_p)
d_p$resid <- residuals(mod)

library(ggplot2)

# Trophic.Level
ggplot(d_p, aes(sample = resid)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ Trophic.Level)

ggplot(d_p, aes(x = resid)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  facet_wrap(~ Trophic.Level)

# Primary.Lifestyle
ggplot(d_p, aes(sample = resid)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ Primary.Lifestyle)

ggplot(d_p, aes(x = resid)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  facet_wrap(~ Primary.Lifestyle)

```
